\chapter{Evaluation}
\label{cha:evaluation}
This section provides an overview of a set of metrics established to compare the final digital product to already established routing solutions for mobile devices, highlight the benefits of the campus app over the manual campus map provided by TU Berlin and measure the improvement of usability when compared to TU Berlin's online web resources. Furthermore, all steps regarding implementation, data acquisition and measurement as well as the resulting consequences for the respective metric are presented.

\section{Campus map verfication}
For campus map verification, there are mainly two different aspects of the digital map that can be evaluated in this process: On the one hand, the coverage of TU Berlin's main campus can be analyzed by comparing the app's campus representation to the representations of other location-based mobile apps. On the other hand, the geocoding capabilities of the system, mainly the mapping of university-specific identifiers and names to their respective campus entities, can be compared to other solutions.

\subsection{Campus coverage}

\subsection{Geocoding capabilities}

\section{Navigation system verification}
The main evaluation point of the navigation system lies in the quality and speed of calculated routes across the campus. To analyze this factor, a comparison between common navigation systems for mobile devices and the campus app is made. This is done by predefining a set of points on the campus, which are further used to serve as start and destination points for multiple route calculations. The provided routes are then compared by walking time and distance.

To define the reference points for route calculation, the existing building data of TU Berlin is used. Each center point of a building is added to the set of reference points and all possible routes (each consisting of one start and one destination building) are calculated. This process is repeated for Google Maps, Apple Maps and the campus app.



\section{Information layer verification}